### 2014-01-09 VerboseNotes.tex
Feynman rules 改訂
git-svn-id: https://dav.misho-web.com/svn/cyghome/trunk/QFT/CheatSheet@1046 c04c31ef-c4f7-4b37-ba94-86ebe2b9c820

%#!platexmake CheatSheet
%%% Time-Stamp: <2014-01-02 02:19:22 misho>
%%% 一部で日本語が使用されています。

@@

>>>>

%#!platexmake CheatSheet
%%% Time-Stamp: <2014-01-09 21:34:25 misho>
%%% 一部で日本語が使用されています。

@@

\subsection{Note on the symmetry factor in the Feynman Rules}
\label{sec:note-symmetry-factor}
If a vertex consists of different particles, the Feynman rule for the vertex is defined as $\ii\lambda$, where $\lambda$ is the coupling constant for the vertex in the Lagrangian.
However, if a particle appears $n$-times in the coupling, we define the vertex as $(n!)\times\ii\lambda$.

This factor $n!$ corresponds to the freedom of contraction, which we encounter when contracting the fields. For example, $\bra{\phi_{k_1}\phi_{k_2}\phi_{k_3}}(\lambda\phi^3\psi)\ket{\psi_p}$ has $3!$ choice of the contraction; thus we define the vertex rule as $(3!)\ii\lambda$, and the amplitude is evaluated as $6\ii\lambda$.

For this multiplication, we have to consider a symmetry factor.
$\begC1{\bra{\phi}}\conC{(\lambda}\endC1{\phi}\begC1{{}^4}\conC{)(\lambda'}\endC1{\phi^3}\begC1{\chi}\conC{)}\endC1{\ket{\chi}}$ has $4\times3!$ freedom, but the rule is already multiplied by $4!\times3!$; thus we have to divide the amplitude by the symmetry factor ($=6$) to obtain $24\ii\lambda\lambda'$.
Similarly, in
$\begC1{\bra{\phi}}\conC{\frac12(\lambda}\endC1{\phi}\begC1{{}^4}\conC{)(\lambda}\endC1{\phi}\begC1{{}^4}\conC{)}\endC1{\ket{\phi}}+\begC2{\bra{\phi}}\conC{\frac12(\lambda}\begC3{\phi}\begC1{{}^4}\conC{)(\lambda}\endC1{\phi}\endC2{{}^4}\conC{)}\endC3{\ket{\phi}}
=
\begC1{\bra{\phi}}\conC{(\lambda}\endC1{\phi}\begC1{{}^4}\conC{)(\lambda}\endC1{\phi}\begC1{{}^4}\conC{)}\endC1{\ket{\phi}}$,
 the symmetry factor $6=(4!)^2/(4\cdot4\cdot3!)$ appears.
(Note that the $1/2$ from the Fourier expansion cancels with the freedom of operators' order.)
